### 2019-05-31 reviews and responses.md
Finished all responses + Discussion improved

Reviews and Authors' Responses for (Busch Khakhalin 2019)
======================

Here you can see peer-reviews for our paper, given in reverse chronological order, and our point-by-point responses to these reviews. Reviews are split into paragraphs, with  responses and comments given after each paragraph.

--------------------------

# Reviews from J Neurophysiol, 2019 Mar 04 (revise and resubmit)

The reviews included Reviewer 2, and Reviwer 3, but no Reviewer 1.

## Reviewer 2

> Busch and Khakhalin use dynamic clamp to generate synaptic conductances of different durations and
then measure the number of action potentials the neurons fire in response to the different
conductances. Using this approach it is found that intrinsic excitability of tectal neurons is temporally
tuned – meaning that they fire best in response to certain durations of synaptic input. This finding is
interesting and unique because typically intrinsic excitability is quantified by generating “input-output”
curves which only vary the amplitude of the synaptic drive, not the duration. Furthermore, the authors
find that various visual stimuli lead to an “adaptive” – not homeostatic – type of temporal tuning of
intrinsic properties such that neurons exposed to shorter stimuli became more responsive to shorter
stimuli and less responsive to longer stimuli., and visa versa. In this study, the effects of visual and
auditory conditioning (and combos thereof) on temporal tuning of both synaptic and intrinsic
excitability are determined. Overall, this study makes great use of dynamic clamp. The interpretation of
the results is limited because the ways in which the conditioning stimuli actually activate tectal neurons
is not empirically established. Clarification, simplification, and consistent use of terms would improve
this paper.

### Minor

> Page 5, line 45/46: “As one possible explanation, auditory and mechanosensory inputs may have
different cellular or subcellular 46 targets in the tectum (Bollmann and Engert, 2009).”
This Bollmann and Engert reference does not address mechanosensory and auditory inputs. This paper
shows that the RGC inputs form a topographic map along the proximal-distal axis of individual tectal
neuron dendrites. This reference should be replaced with an appropriate reference.

We replaced this sentence with the following sentence: "This was not necessarily surprising, as auditory and mechanosensory inputs are known to have different cellular targets in the tectum (Pratt and Aizenman, 2009; Felch et al., 2016; Truszkowski et al., 2017), and thus can recruit tectal excitatory and inhibitory networks to a different extent."

> Also, how can having different subcellular targets could explain the results? This concept needs to be
clarified – an example would be helpful.

We now added the following sentence to the text: "Alternatively, acoustic stimuli may be inherently weaker in triggering plasticity effects in tectal neurons, as compared to visual inputs, they arrive at different subcellular compartments within the dendritic tree (Hiramoto and Cline, 2009; Deeg et al., 2009), which can affect their influence on neuronal plasticity (Richards and Lillicrap, 2019)."

The actual rationale here is that inputs to different parts of the dentritic tree can trigger action potentials of different shape, and differentially affect calcium dynamics in the neuron, which, together with non-linear processing in dendrites may make different types of inputs unequal in terms of their effects on neuronal activity at one hand, and their impact on their tuning on the other. Two extreme examples of this effect would be the impact of climbing fibers on the activity of Purkinje cells in the cerebellum, and inputs to the apical dendrites of pyramidal cells in the cortex. However, we feel that this topic is really deep, yet somewhat tangential to the main message of the paper, so we tried to limit our exposition to this short sentence with a reference.

> Page 13 line 46: The authors use the term “distributed sensory network” . Is a “distributed” network a
special type of sensory network?

Technically, even two neurons form a network (a true mathematician would probably argue that even one neuron forms a network!), so we tried to remind the reader that the tectal network is a relatively large one, with broad recurrent excitation, and in which different parts of the sensory input are processed in parallel. Thus we used the word "distributed", as in "distributed computations", as opposed to more local and covergent networks, such as, for example, the Mauthner cell network in the hindbrain. If you think that this word is not appropriate here, please let us know, and we'll remove it.

> Page 3, line 26,27,28: “In the tadpole tectum, looming stimuli are known to elicit stronger responses
compared to instantaneous inversions, yet the suppression yielded by looming stimuli was weaker”.
The logic of this sentence is not clear. Why exactly would the authors not expect that the looming
stimuli that is known to elicit stronger (synaptic?) responses would also cause less suppression (of action
potentials?) than does flashes?

This was unexpected as intrinsic plasticity is typically described as a homeostatic mechanism, and thus a stronger stimulus should, in principle, cause stronger suppression of spiking. To make it more clear, we now changed this sentence:

"This was likewise unexpected, as compared to flashes, looming stimuli are known to elicit stronger tectal responses (Khakhalin et al., 2014; Khakhalin, 2019), yet their effect on neuronal excitability was weaker, seemingly violating the principle of homeostatic plasticity (Pratt and Aizenman, 2007; Turrigiano, 2007)."

### Major

> Page 5, line 2,3,4 : But do the authors actually know how/if their conditioning stimuli is altering the
pattern/strength of synaptic input? If so, this data should be shown. If not, it should be made clear that
this is an assumption/ theoretical prediction.

In the submitted document, these lines said:

"We then examined whether different types of sensory activation would differentially reshape  
temporal intrinsic tuning in tectal neurons. As changes in intrinsic properties seemed  
homeostatic (increased activation led to reduced spiking), by the same logic, one could
expect shorter stimuli (flashes) to selectively suppress responses to shorter synaptic inputs."

Here we are not talking about our conditional stimuli alterning the pattern of synaptic inputs; we just refer to the fact that during the stimulation, shorter stimuli (such as instantaneous flashes, that are technically the shortest stimuli possible) would cause faster visual inputs to the tectum, compared to longer stimuli (second-long continuous transitions) that would cause slower inputs. In this study, we did not measure actual synaptic responses in the tectum during conditioning, but from previous studies we know that responses to flashes are actually much shorter than responses to looming stimuli (Khakhalin 2014), as described in the beginning of the "Results" section, and then again in the "Methods".

We now changed the wording of this sentence, to make this logic clearer.

"As changes in intrinsic properties seemed homeostatic (increased activation led to reduced spiking), by the same logic, one could expect that exposure to shorter sensory stimuli (flashes; Khakhalin et al. 2014) would selectively suppress spiking in repsonse to shorter synaptic inputs."

> Page 5, line 5: “As changes in intrinsic properties seemed homeostatic (increased activation led to
reduced spiking), by the same logic, one could expect shorter stimuli (flashes) to selectively suppress
responses (action potential firing or synaptic responses or both?) to shorter synaptic inputs”.
What intrinsic properties were measured? It appears that only number of spikes was counted. The
authors report that long-term increase in activation (by visual stimuli) decreased spiking – but was that
due to a decrease in intrinsic excitability or because the stimuli decreased synaptic strengths – which in
turn led to the observed decrease in number of spikes fired? This is critical to define.
Second, do the authors mean to say that shorter stimuli would actually (somehow) suppress responses?
How would that work?

All our measurements of the number of spikes relied on the dynamic clamp experiments, that used simulated synaptic inputs, and not real synaptic stimulation. It means that any changes in spiking we observed in the previous section of the results is necessarily coming from the change in intrinsic properties. While a change in the synaptic strength in response to sensory stimulation is possible, and indeed is investigated below in this paper, it would not have affected the number of spikes we reported in figures 1 and 2, as they were recorded in dynamic clamp experiments.

As described above, to make this point more clear, we now reworded this sentence:

"As changes in intrinsic properties seemed homeostatic (increased activation led to reduced spiking), by the same logic, one could expect that exposure to shorter sensory stimuli (flashes; Khakhalin et al. 2014) would selectively suppress spiking in repsonse to shorter synaptic inputs."

> Page 5 line 15: Does the word “preferred” (in quotes) mean that the cells fired more action potentials?
Maybe refrain from personification.

We hope that the word "prefer" is intuitive enough to most readers, and it is also way shorter than "responded with more spikes", or "produced more action potentials", which allows for easier writing, and hopefully better reading. Realistically, we believe that most readers of this article will not be at risk of truly personifying neurons, so while we take care to never personify animals, we are intentionally a bit more relaxed when writing about cells, channels, receptors etc. Still, to make this sentence better, we removed the quotes around "preferred", to make it more obvious that it is not an antropomorphic allegory, but just a normal word use. Overall, we use the word "prefer" 6 times in the paper, always in the same way, when comparing spiking responses to different stimuli.

> Bottom of page 5, “Effects of acoustic and multisensory stimulation:

> - Is the Auditory stim conditioning referred to as “click sounds” the same as the “startle-inducing
sounds” that is mentioned in the very next sentence? If these are indeed the same stimulus it should
be referred to by the same name, consistently. If they are not the same stimuli, the differences need to
be explained.

These are the same stimuli, and we changes these two sentences to better reflect that:

"To test this question, we exposed tadpoles to four hours of behaviorally salient, startle-inducing “click” sounds, provided at the same frequency (every second) as visual stimuli in the first set of experiments. We found (Fig 1D, E) that exposure to these sounds..."

> - Tadpoles are exposed to 4 hours of click sounds and then intrinsic tuning was measured. No difference
in spiking was observed. This was interpreted to mean that acoustic stimuli did “not strongly activate
tectal circuits during conditioning”. How can the authors know this? Instead, why not determine this
directly by recording click-evoked responses from tectal neurons? Or, if tectal cell responses to auditory
stimuli have been described in previous work, this should be included here.

Unfortunately, recording of tectal reponses to sounds in tadpoles does not seem to be possible at present, not just in our lab (where we don't posess the technical capability to perform in-vivo experiments), but in most labs. Technically, the problem amounts to two limitations: one, to record from a tadpole, one has to paralyze it with a nicotinic receptor anatagonist, which seems to disrupt hearing. Second, even if this limitation is overcome, the sounds has to be strong enough to stimulate the inner ear, but weak enough not to disrupt a recording. Conceivably, both challenges may be overcocme with the use of high speed calcum imaging in freely moving tadpoles, but this is also challenging, and to our knowledge no lab has attempted these recordings yet.

To show that our statement is just a conjecture, we replaced "suggets" with "may suggest" in this sentence: "This may suggest that acoustic stimuli did not activate tectal circuits strongly enough during conditioning..."

> Page 6 line 24: Remove the phrase “are not easily interpretable”. Instead, please interpret your data
clearly.

We removed this phrase from the main text.

> Page 9: The authors report that the synaptic data show that “prolonged sensory activation with short,
frequent stimuli reshaped synaptic transmission in the tectum, making it faster through selective
potentiation of visual inputs from the eye”.
How do the authors know the effect is selective for RGC inputs? It appears that they only measured
RGC-evoked responses. Auditory/mechanosensory -evoked responses were not reported.

Stimulation in the optic chiam evokes two types of synaptic responses in tectal cells: direct short-latency responses, and delayed recurrent inputs from other cells in the tectum. By looking at average currents early and later after the chiasm shock, we can independently assess the contribution of direct and recurrent inputs. In the paper, in a paragraph before the sentence in question, we report that after stimulation, direct current increased in amplitude, while recurrent currents remained unchanged. This is what we call selective potentiation of direct inputs, as opposed to recurrent inputs. We do not claim that mechanosensory inputs did not change, as we have not recorded auditory responses in this study (recordings of auditory responses are considerably more difficult; see Felch 2016).

To make sure that we are not misunderstood by the reader, we replaced the original sentence with the following two sentences: "This shows that prolonged sensory activation with short, frequent stimuli reshaped synaptic transmission in the tectum, making it faster through selective potentiation of visual inputs from the eye, compared to recurrent inputs within the tectum that remained unchanged. In this study, we did not perform a similar analysis for acoustic projections from the hindbrain."

> Page 9 line 9: “cells exposed to stronger stimuli preferred shorter synaptic inputs in dynamic clamp…”
Here, (where “preferred” is not in quotes) does “preferred” mean to fire more action potentials?

Yes, as we answered above, we now don't use quotation marks for any of the words that aid writing and understanding, so "preferred" consistently means "responded preferably", or "produced stronger responses".

> Page 10 (last paragraph): Were action potential thresholds measured? Were there any differences in
this parameter across groups?

We did not measure action potential thresholds, as we knew from the previous study (Ciarleglio Khakhalin 2015) that potential threshold measurements in tadpole tectal neurons are unreliable, in the sence that they strongly depend on the current that is used to evoke the spikes; this is almost certainly due to rapid Na channels inactivation.

> Page 10 line 8: The “correction for position” is difficult to understand here and in Methods. Why was a
linear regression used – and how is it determined to generate an accurate “correction”? How do the
data look without this correction?

The linear regression is used because it is the simplest way to compensate for a confounding variable (see ANCOVA, or linear models with confounding variables in general). Linear models are the "default" approach in statistics, and are typically used when the exact relation between two variables is unknown, or when the amount of noise is so high that practically there is no difference between linear and conceivable curvilinear regressions. Both situations apply here.

There is no way for us to determine whether our adjustment is "accurate", as there is no experimental way to make more caudal (and thus younger) cells in the tectum older. This situation is not unique however, as it arises in any study that uses linear models, in any field, from ecology to social sciences. When running a study in a human population, if a variable of interest depends on age, one would typically adjust for age by including age as a covariate in the linear model. Here we perform the same standard procedure. After rereading the manuscript, we realized that we used terms "adjustment", "correction", and, in one case, "compensation" interchangeably, which might have led to some confusion. To stress that we are talking about a basic statistical technique of covarate elimination, we now consistently use the term "adjustment" throughout the manuscript.

In the paper, we do not report how the data would have looked without this adjustment for covariates, as it would not be statistically sound to run a linear model with a known covariate without including it. In most cases, it may be expected to underestimate effects for analyses across groups (as the effect of the covarate would be included in the error), and overestimate effects for explanatory linear models in the second half of the paper (as the covariate would contribute to both the dependent, and the independent variables in the model). For each particular comparison, however, both d, F and p values would differ by a small random value. For example, when the measure of "spikness" (one of the three tuning parameters) is compared across groups, a correct analysis with adjustment for position (`smean~rostral+medial+group`) produces a p-value of 0.01. Without adjustment (`smean~group`) we get a p-value of 0.03. For temporal tuning, correct analysis (`sbend~rostral+medial+group`) gives p=0.0304, and if covariate is left unadjusted (`sbend~group`) we get p=0.029. Similarly, for amplitude tuning, we get p=0.00209 for correct analysis, and p=0.0024 for unadjusted analysis. For the analyses in which temporal tuning and ampliltude coefficients were explained through intrinsic parameters, the optimal models were not qualitatively different between adjusted and unadjusted data, but all values in the table were slightly different.

If you would like us to numerically re-run any part of the analysis without a position adjustment, please let us know which part you would like to see. All scripts that produce technical checks, visualizations, and analyses related to this adjustment are available for inspection at: 
https://github.com/khakhalin/Dynamic-clamp-2018/blob/master/dynamic_other_vars.R

> Page 13, line 7: What does “intrinsic spikiness” mean? Is this another term for intrinsic excitability, or is
this a different parameter? In fact, throughout the manuscript, there are many terms including “intrinsic
excitability, spikiness, number of spikes, and intrinsic spikiness that probably mean the same thing – or
maybe not. If these terms all mean the same thing, then one term should be chosen and used
throughout. If they all refer to different entities, then each one should be clearly defined.

In this updated version, we consistently use "spikiness" as a name for the variable that is for the first time introduced in Figure 2C. For earlier results section, we replaced "spikiness" with "average number of spikes", and in all instances when we talked about the general idea of neuronal excitability, we called it neuronal excitability. As the only exception to this rule, we opted for the word "spikiness" in the first paragraph of Introduction, to prime the reader for the use of this word later in the text.

The expression "Intrinsic spikiness" is no longer used in the text.

The expression "Intrinsic excitability" is now used only as a name for the general phenomenon, that technically may be probed with different alternative protocols and measures. In this  study, we mostly assess it using a more narrowly defined "spikness" variable, and then, towards the end, using a less informative measure of the number of spikes observed in current clamp experiments. In the text, we now consistely use "intrinsic excitability" when we make inferences about the actual properties of the neuron, and use "spikiness" when we talk about the actual value we calculated to assess and quantify intrinsic excitability.

> Page 15 line 34: The average membrane resistance (330 MOhm) seems a bit low for these small immature
tectal neurons. Were the neurons healthy? What was the average resting membrane potential? Were
there any set criteria for Rm, Vm, or access resistance?

Thank you for catching this error! We had an order of magnitude typo here, and the average membrane resistance was in fact 3.2±1.5 GOhm. This value is a bit higher than reported earlier (1.2±0.8 GOhm in Ciarleglio Khakhalin 2015; about 1.1 in Khakhalin 2011), but within what can be reasonably expected for healthy neurons of this size. To make sure that this number reflects our observations, for the purpose of this response, we verified this value by reanalyzing our data offline, and obtained a similar estimation for the average membrane resistance (2.9±1.4 GOhm). We have several reasons to believe that the cells included in this study were healthy. First, the access resistances obtained were good (30±14 MOhm, which is similar to benchmark 50±23 in Ciarleglio Khakhalin 2015, and about 20-30 MOhm in Khakhalin 2011), indicating that higher membrane resistance is not likely to be due to poor access. Second, our neurons reliably produced spikes in current clamp mode, and active currents in response to membrane depolarization in voltage clamp mode, and the quantification of these active conductances yielded numbers similar to those reported before (Ciarleglio Khakhalin 2015). The quality of voltage clamp, as judged by the absence of escaping active currents in IV protocols, was good in the majority of neurons (82% of all neurons recorded). Finally, a few cells that had their membrane resistance change drastically during the recording session, were discarded at the data collection stage. We did not measure the resting potential directly, but measured the holding current necessary to bring the cell to -60 mV in voltage clamp mode, and the mean value of this I_hold value was -8.2±4.8 pA, which is not higher tha what was previously recorded in the literature (19.8±16.6 in Ciearleglio Khakhalin 2015).

Overall, these arguments make us believe that the quality of our patches was good enough to draw the conclusions reported in the study. Note that all data for all cells is available at the following address:
https://github.com/khakhalin/Dynamic-clamp-2018/blob/master/Data/data_mainInput.txt
In this dataset, groups "Naive", "SlowF", and "SlowC" should be ignored, as they were not included in the study. Note also that github doesn't show columns correctly, so if you want to inspect the data, we recommend that you download it and open it in a spreadsheet.

As the mistake your noticed (thank you again for being so thorough!) was a typo in the text, and not a mistake in the data, it does not affect any other calculations reported in this paper.

> Figure 1 D and E: The average number of spikes fired as a function of conductance duration and all at 3
different peak conductances. Are there any trends at the single cell level? For example, if a given
control neuron fires relatively few spikes in response to 500 ms duration, does it also tend to fire
relatively fewer spikes with 1000ms duration? Or maybe if a given neuron fires a relatively high number
of spikes with 500ms then it will typically fire less at 1000ms duration? Although the individual
responses are shown in 1D, it is impossible to follow the number of spikes fired by any individual neuron
in response to the different conductance durations.

At least for high-spiking neurons, the answer to this question can actually be inferred from Figure 1D: those curves that are placed high on the graph for duration of 500 ms, also tend to remain high for 1000 ms, indicating that high-spiking neurons tend to be high-spiking across the range of durations. Numerically, the correlation coefficient between the average number of spikes in response to 500 and 1000 ms, across all neurons, is r=0.94. Arguably, this was to be expected as intrinsic excitability of neurons (whether the neuron is spiky or not) is a much stronger feature than the temporal tuning. (Temporal tuning in the tectum was not even described until this paper, while the variation in intrinsic excitability in the tectum is well known for more than 20 years).

> Figure 4 D, E and F: Is “spikiness” synonymous with “N Spikes”? If these terms do not mean the same
thing, they should be defined. If they do mean the same thing, then one term should be chosen and
used consistently throughout.

As described above, "spikiness" is now consistently used across the manuscript as the name of one of three variables used to quantify intrinsic tuning of tectal neurons.

## Reviewer 3

> Busch and Khakhalin use the Xenopus tadpole tectal model to address the question of
how neurons modify their intrinsic properties when an animal experiences different
types of sensory input. Animals that underwent different forms of sensory experience
had tectal neurons that decreased their overall excitability. This is consistent with the
idea of homeostatic regulation of intrinsic properties. Surprisingly, when the sensory
experience involved short pulses of light, neurons because more responsive to short
inputs even though there was an overall decrease in spiking. This adds a layer of
complexity to the idea of homeostatic control of a firing rate set point because it shows
neurons can independently regulate total spiking in the opposite direction from their
responses to synaptic inputs of different durations. This is a really interesting idea. The
data are sound and well-analyzed (beautiful figures, open source data, etc.). However,
from top to bottom, this manuscript was incredibly difficult to read. Due to organizational
issues, lack of a strong rationale, and inconsistent/confusing wording, I’m not entirely
sure I understand the main message the authors are trying to convey. My summary is
largely based on what I was able to pick out of the figures. I strongly suggest the
authors perform a massive rewrite and really emphasize the main points.

### Major Comments

> Abstract:

> It’s hard to follow what was actually done and found in the paper based on the abstract.
The abstract starts off with a statement about homeostatic regulation through changes
in intrinsic properties and then gives a detailed account of what was found without any
context. It would be easier to assimilate the information if the abstract follows a more
standard format: start with the area the paper clearly contributes to, build the rationale
for the study, state the hypothesis (or driving question), describe how this was tested
and what was found, and say why it’s important.

We rewrote the abstract, stressing the big question to which the paper contributes (principles of tuning and learning in biological neural networks), and the main message of the paper. We also made the account of our findings a bit less detailed.

> Introduction:

> I feel the same way about the introduction as I do the abstract; as written, it doesn’t give
the reader a great sense of what’s to come. If the setup for the paper is that
“connectivity isn’t everything— intrinsic properties matter,” this is a good start, but it
could be made much clearer. Further, the questions of the study are not presented until
the third paragraph, and there is no rationale for these questions. I suggest the authors
completely revamp the introduction in way that guides the reader from
regulation/plasticity of intrinsic properties (the big picture question) to why there’s a
reason to think neurons in their system can tune their responsiveness to different kinds
of inputs they receive (a specific hypothesis). This is what the authors actually address,
but they really don’t provide much of a rationale for their specific experiments, which
weakens the paper.

We now made the introduction clearer, and better organized, going from the big problem, to specific questions. We moved the presentation of the questions for the study to the very beginning of the second paragraph, and provided a rationale for these questions in the first paragraph.

> Results and Discussion:

> - The rationale for the use of dynamic clamp and the parameters chosen are not clear.

Previously a detailed justification for the use of dynamic clamp was only given in the Discussion section; we now moved it to the end of Introduction.

> Is the dynamic clamp current injection protocol mimicking the physiological synaptic
input? This should be stated.

We updated the next sentence, by adding a reference to it: "We used conductances of 4 different durations (100, 200, 500, and 1000 ms), and 3 different amplitudes (peak conductances of 0.2, 0.5, and 1.0 nS), matching the range of synaptic currents observed in tectal circuits in vivo (Xu et al. 2011; Khakhalin et al. 2014; Ciarleglio et al. 2015; see Methods)."

> - I think the experiment in this study is pretty straightforward, but the description of the
results is not. As far as I can tell, tadpoles were exposed to different stimuli (flash,
looming, multisensory, etc.). Then, tectal neurons were stimulated using either 1)
dynamic clamp or 2) more physiologically through optic chiasm. This allowed the
investigators to assess firing when they changed the stimulus intensity and stimulus
duration of the dynamic clamp or the synaptic response to presynaptic stimulation.
They then looked how the different types of sensory experience influenced firing
responses. The rest is data analysis that looks at how the duration, amplitude, total
spiking, and synaptic input is changed in different sensory contexts and sensory
stimuli might interact. I think the text can be made much simpler so the basic design
is easier to grasp. I suggest thinking about what the take home point is for each
figure and craft the results section around those messages.

Thank you; we tried to follow this advice in this revised version of the manuscript!

> - There is also an inconsistent use terminology. For example, “spikiness” and
“excitability” are sometimes used in attempt to describe distinct cell characteristics,
but sometimes it seems like they are being used interchangeably. This makes
interpreting the written results difficult to follow.

As described in our responses to Reviewer 2, in this updated version, we consistently use "spikiness" as a name for the variable, for the first time introduced in Figure 2C. For earlier results section, we replaced "spikiness" with "average number of spikes", and in all instances when we talked about the general idea of neuronal excitability, we called it neuronal excitability. As the only exception to this rule, we opted for the word "spikiness" in the first paragraph of Introduction, to prime the reader for the use of this word later in the text.

> - “Compared to control, neurons from animals exposed to visual stimuli had a flatter
amplitude tuning curve (Fig 1D,E).” What is the reader supposed to be looking at in
the figure? Is this referring to larger conductances for a fixed duration? I think this is
what the authors’ mean, but more intuitive descriptions that matches the figures or
vice versa is needed to make the whole results section easier to digest.

We improved the sentence:

Compared to control, neurons from animals exposed to visual stimuli had a flatter amplitude tuning curve (**less of a difference between columns of Figure 1D, or between curves in each individual panel of Figure 1E**), as they did not increase their spiking as fast in response to larger conductances

> - Excessive wordiness makes the results difficult to read and interpret. For example,
on pg. 9, 4 paragraphs (411 words) are used to describe two graphs, each showing
a different linear relationship (Fig. 3E-F). Why not simply say, “A positive relationship
between temporal tuning and synaptic duration implies a co-tuning of preferr
parameters.”

Of these four paragraphs, the last one reports a negative result that is not shown in Figure 3. The first three paragraphs are indeed dedicated to panels Figure 3EF, but we cannot in good faith replace them with one sentence, as the statement we are making is not statistically obvious. We believe that an attempt to gloss it over by pushing the statistical analysis to the methods section would not be fair. The issue, as we describe in the text, is that while graphs 3E and 3F have similar axes, and both show a linear dependency, they have opposite slopes. Had the slopes be the same, we would have just reported a correlation in one sentence. But with the data we have, we need to be more specific, as we observed that the within-group correlation is positive, yet the between-group correlation is negative: a situation that is typical for larger datasets, but that some readers may not be familiar with.

We are aware that this section is read differently by different readers, and we received some conflicting advice in the past about the level of details needed to communicate this message. For example, those who have explicitly discussed Simpson's paradox in their statistics classes typically find our use of this term very helpful, while those who haven't used this term before, may find it distracting. For now, we opt to include it, even though it seems to be appreciated by only one half of our target audience.

Nevertheless, in current version of the paper, we shortened this section from 411 to 294 words.

> - In reference to the text at the bottom of page10: Is it possible to include a summary
figure of data from the stepwise generalized linear model in Fig. 4? The purpose of
this part of the study was to try and identify electrophysiological mechanism(s) that
explain the tuning variables. The heart of Fig. 4 should be that they don’t, and figure
that captures this would be helpful.

We are not sure how this fact can be illustrated. We searched for papers that use stepwise linear models, and it seems they do not usually include summary figures to describe these models. (They often include tables, listing R-squared, F, df, and p-values for every component, but as in this paper we are not particularly concerned by the precise composition of our models, we doubt a table like that would be useful here). Do you have any specific type of illustration in mind, in which case, could you provide a reference, or maybe describe it? The only type of illustration we could come up with is a scatterplot that compares values predicted from the model to actual values, but it is already shown in the paper, as Figure 4E.

> - If these data are interpreted as homeostatic (neurons became less excitable
following sensory experience), is there any evidence that tectal neurons have
different firing activities during the different kinds of sensory experience? For
example, does sound engage tectal neurons less than visual stimuli, and might this
explain some of the trends in the data?

This may indeed explain some of the trends in the data, but unfortately, as we described above, recording of tadpole tectal reponses to sounds does not seem to be possible in our lab, and is fundamentally technically problematic in principle. We can only guess that stimuli so different as flashes, looms, and sound clicks, would produce different responses in the tectum, as their dynamics is drastically different, as they engage different circuits before they rich the tectum (retinal for visual stimuli, hindbrain for acoustic), which undoubtedly shapes the dynamics of tectal activation (based on Khakhalin 2014, Felch 2016, Truszkowski 2017, Khakhalin 2019).

We now make this point more clear by referencing it in the beginning of the paper (as described above), and then again, at the very end of Discussion, by describing how our observations can be used to predict actual, and yet unobserved, activation of the tectum during multisensory stimulation in vivo:

"Finally, based on the multisensory phenomena reported in this paper, we predict that even though multisensory stimulation tends to increase tectal responses in vitro (Felch et al., 2016; Truszkowski et al., 2017), it would be likely to reduce peak activation in vivo."

&nbsp;

--------------------------

# Reviews from J Neurosci, 2018 Nov 01 (rejection)

These reviews were received on Nov 1st 2018 from J Neurosci, as this paper was rejected by the editor. We decided to include these reviews to our next submission, and to this repository, and provide our responses to each of the critiques.

## Reviewer #1

> The study is full of clever ideas and technical expertise, but it is too phenomenological and descriptive at this point, lacking a systematic step by step dissection of the underlying mechanisms. The hypothesis is fascinating but it isn't really tested directly.

> This paper takes advantage of the Xenopus laevis tadpole optic tectum as a model system for examining how prolonged (4h) exposure to specific forms of conditioning sensory input to a neuron alter its intrinsic excitability and tuning for firing action potentials. Using the dynamic clamp method the authors show that cells exposed to intensive (presumably brief) visual (but not other sensory) stimulation shift their preferred input to favor shorter-lasting synaptic inputs, which the authors describe as "adaptive" plasticity that allows the cell to fire in response to the stimuli which it recently experienced. They also discovered a so-called homeostatic plasticity within conditioning groups such that those cells that have brief rapid responses to optic chiasm stimulation tended to prefer long slow inputs whereas cells with slower repsonses seemed to be tuned to brief inputs. None of these properties correlated with position in the tectum and therefore with presumptive cellular maturation.

> Although the central question and ideas that make up this paper are very interesting and potentially constitute a very new way of thinking about plasticity of intrinsic excitability, the paper is almost entirely descriptive and phenomenological, with a number of holistic inferences that, although thought provoking, are not necessary covered withing the scope of the experiments presented. For example, the idea that adaptive vs. homeostatic mechanisms are at play infers too much of a desired outcome for the system without any real justification for such an attribution, especially without having characterized the actual network responses to the conditioning stimuli used. Maybe the plasticity doesn't serve the purpose of making the cells more sensitive or regulating dynamic range. Perhaps it does something useful but unanticipated for the network as it functions in the entire animal. We don't know and shouldn't presume. Also there were numerous points where the data presented here disagree with previous publications by the senior author and although some handwaving explanations were provided, no actual experimental evidence for these differences was presented.

> A reasonable attempt was made explain the conditioning changes observed at a more mechanistic level, but at least for the dynamic clamp experiments the authors were unable to account for their observations based on traditional mechanisms for regulation of intrinsic excitability. Overall the paper is fascinating and introduces a potentially interesting new way of thinking about plasticity, but doesn't quite provide enough of a mechanistic underpinning nor of a functional context for the model. It comes across as a slightly contrived set of systematic observations. The paper has a few minor technical issues, but suffers mainly from failing to systematically test any hypothesis in a satisfying and reductionist way. The paper would likely be better suited either for a more specialized cellular neurophysiology journal or, if the narrative can be tightened up to clearly convey its novelty, for a short format high-impact report.

&nbsp;

We are grateful for the kind words, and agree with the assessment that our paper is largely phenomenological. Unfortunatelly, in a small teaching-oriented college, we simply cannot run most experiments needed to dissect the mechanisms of intrinsic temporal tuning. We cannot perform in-vivo experiments with sensory stimulation; and cannot measure or change gene expression in isolated tectal cells. Even a detailed electrophysiological study of isolated ionic currents would have taken extra 2-3 years, so we decided to publish this phenomenological report now, hoping that the very phenomenon we describe is interesting enough to be followed up by other labs, and potentially in different models.

In this, updated version of the paper, we try to be more explicit about possible next steps that may be taken to study the mechanisms behind the exciting phenomenon we discovered.

### Specific points:

> 1.	Page 2, Line 29 the word "thorough" is not appropriate here. If the search had been thorough, an answer would have been found. Instead "extensive" may be more appropriate.

Corrected.

> 2.	Page 2-3, The dynamic clamp is not very commonly used. It would be useful to provide more detail about how it differs from simple step injections of current and why this might matter.

Thank you for reminding us about that! We now added a one-sentence description of dynamic clamp into the last paragraph of the intro section:

"Unlike for more common voltage and current clamp techniques, in dynamic clamp, the electric current injected into the cell is dynamically adjusted, based on a predefined formula that depends on cell membrane potential and time."

> 3.	Page 3 line 18, "when tadpoles were exposed to" -> "in tadpoles that had been exposed to"

Corrected.

> 4.	The statistical tests applied throughout the paper are unclear and need more explanation. The number of degrees of freedom (e.g, 667) seems far too high, suggesting that multiple recordings from the same cell under different conditions are being treated as independent N. However recordings from the same cell under different temporal or amplitude conditions are far from independent measurements. It seems that a two-way or even three-way ANOVA might be the best approach for examining the variables of conditioning stimulus and response property.

The number of degrees of freedom in question comes from a full linear model, and does not mean that we treated different conditions as independent. Typically, to compare the numer of spikes in two conditions we used a 4-way ANCOVA (injection amplitude and duration as ordinal variables; group id as a nominal variable) with interactions, and repeated measures (cell id was included as a nominal variable). For example, for a comparison of "Control" group to "Sound" group we used the following R code (see `dynamic_spikes_plot.R` for the full code):

```R
ds = subset(d,is.element(Group,c("Control","Sound")))
summary(aov(data=ds,Spikes~Shape*Amp + Shape*Group + Amp*Group + Cell*Shape*Amp))
```

In addition to a full description of our analysis approach, as given in the "Methods", we now included a second, shoter description of it in the "Results" section, to make the paper more readable. We also improved the description in the "Methods" section; and added several extra referrals to the "Methods" section from the "Results" section, to remind the reader that extra information is available there. In particular, in the sentence that reports the first F-value for this study, we now added the following commentary:

"F(1,677)=30.4, p=5e-8, n cells=28, 29. Here and below, when F-values are reported, we use 4-way fixed effects analysis of variance with selected interactions, and cell id included as repeated measures factor, see "Methods" for a detailed description."

> 5.	Page 5 line 19 "This suggests that unlike the change in overall intrinsic excitability, the average temporal retuning was "adaptive" rather than "homeostatic", as neurons exposed to shorter stimuli (flashes) became more equipped to process shorter activation patterns." While it is appealing to give a name to this form of plasticity, the looming stimulus, which can be thought of as slower than the flash nonetheless caused a similar type of shift in response preference in the neurons, which runs counter the notion that there exists a meaningful dichotomy of adaptive vs. homeostatic change here.

We think that there exists a certain dichotomy here, but we agree that our description was not clear enough. We now replaced it with the following, longer exposition:

"... This suggests that the change in overall intrinsic excitability, and the change in temporal tuning, follow two different kinds of logic. The overall excitability is homeostatic, as neurons became less excitable in response to stronger stimulation. The temporal retuning however can be better described as "adaptive", as neurons exposed to shorter stimuli (flashes) became relatively *more* responsive to shorter stimuli, and less responsive to longer stimuli, which is opposite of what one would expect for a purely homeostatic retuning. We chose to call this type of plasticity "adaptive", as presumably it means that after exposure to faster stimuli, neurons become more equipped to process faster patterns of activation (Stemmler and Koch, 1999)."

> 6.	Figure 2B Amplitude is misspelled.

Fixed.

> 7.	The narrative logic on page 9 lines 11-22 is confusing. How does homeostatic plasticity, if it truly is at play here, "increase their ability to respond to unusual patterns of synaptic activation." If anything wouldn't homeostasis decrease dynamic range by pushing all cells toward a central set point?

We agree that the word "homeostatic" was confusing in this context. We now simplified the sentence:

"This means that individual neurons tended to tune their intrinsic properties away from the typical statistics of their inputs, decreasing their responses to common input patters, and enhancing responses to unusual patterns of synaptic activation."

> 8.	Page 12 the section on position within the tectum feels like it was just tacked onto the end of the paper. This should be integrated earlier in the paper perhaps to go with figure 1A.

We placed the decription of this effect at the very end of the paper to downplay it a bit, as some aspects of this effect were described previously (Khakhalin 2012; Hamodi 2012; Ciarleglio 2015), and also we don't want to distract the reader from our other findings. At the same time, it is important to mention that we observed a strong correlation with position: first, because it provides an independent replication of previously observed patterns, and second, as the dependency of cellular properties on position necessitated an adjustment for position during analysis (see "Methods"), and this analysis decision needed to be justified. The implication for the principle of parameter degeneracy in development (Prinz 2004) is also interesting, and we wanted to point at it, even though we don't have enough material in our paper to fully explore this theme.

## Reviewer #2

> There is indeed some merit in the goals of the study (to see if adaptation changes with experience in addition to conductance/response gains), but it has so many technical problems at present that it would need to be radically different (perhaps with experiments and analysis re-done) to perhaps be of interest to the broad readership of J Neuroscience.

> The manuscript by Khakhalin and Busch describes the impact of experience on the intrinsic response properties of tectal neurons in the developing tadpole. The authors describe a surprising phenomenon: not only is the response gain of the conductance response altered by experience (in a homeostatic manner, if we imagine that the experience produces increased activity, which isn't shown), but the adaptation of the response is also impacted. The authors then go on to make some arguments, in my opinion unconvincing, that properties of the duration of the polysynaptic response to an optic chiasm shock are related to these intrinsic properties.

> The paper is, in my opinion, hard to read because the authors are often not clear about what they mean, refer to previous parts of the paper assuming the reader has a perfect memory or is willing to turn back and forth across pages, and use non-traditional terminology. In addition, the statistically analysis appears to be pooling many data points from non-independent cells, rendering many of the conclusions invalid. Further, the experiments intended to measure intrinsic properties (in Figures 1 and 2) appear to have been performed without synaptic blockers, so the the adaptation that they measure could be in the circuit (inhibitory recruitment) instead of in the intrinsic properties.

> There is indeed some merit in the goals of the study (to see if adaptation changes with experience in addition to conductance/response gains), but it has so many technical problems at present that it would need to be radically different (perhaps with experiments and analysis re-done) to perhaps be of interest to the broad readership of J Neuroscience.

&nbsp;

On excessive cross-references between different parts of the paper: thank you for pointing it out! We now tried to link different parts of the paper more tightly together, by gently reminding the reader about previously made statements every time we make a reference.

On statistical analysis: we do not pool data points from non-independent cells, but used a repeated measures analysis, verified by mixed model analysis (see Methods, as well as corrections described above, in our responses to Reviewer 1).

On synaptic blockers: it is true that we measured intrinsic properties without synaptic blockers, but this is considered a standard practice in this preparation, as most earlier studies measured IV curves without synaptic blockers (Azenman .. Cline 2002; Aizenman .. Cline 2003; James .. Aizenman 2015; Ciarleglio Khakhalin .. Aizenman 2015). What makes this practice possible is a relatively weak recruitment of polysynaptic activity after activation of a single tectal cell (Bell .. Aizenman 2011).

### Specific points:

> Introduction:

> Page 3, line 23: Sentences that begin "First" and "Second" need subjects

Fixed.

> Page 3, line 34: "spikiness" -> not a precise word. "Changes in f-I curve", perhaps? Or "Average firing rate"?

We think that this word reflects the nature of the property we are describing rather well (the property of being easily excitable, and able to generate spikes), and we also define it within the paper itself. We prefer it to other alternatives, as they are either too general ("intrinsic excitability") or too bulky ("ability to produce spikes in response to stimulation").

Of the two options offered by the reviewer, the first one ("Changes in f-I curve") is quite long, and also does not reflect the nature of our measure, as with it we are trying to assess average ability to produce spikes, not a change of any sort. The second one sounds good and standard, but would actually be misleading in this particular case, as the term "Avereage firing rate" is defined as the average rate at which a neuron fires in the network; usually either in the context of in-vivo recordings, or in-vitro spontaneous activity. Accordingly, a true "Average firing rate" for our neurons would depend both on their intrinsic properties (the measure we call "spikness here"), *and* the strength of their synaptic inputs, as well as their position within the network (centrality measure). In short, while "Average firing rate" is an important property, we did not address it in this study, and had no opportunity to observe a proxy for it.

> Page 3: "traces left by multi sensory stimuli in the tectum of freely behaving tadpoles" -> unclear what is meant by the this sentence until later; suggest re-word

Reworded: "for the first time we were able to look at tectal network retuning in response to multisensory stimuli in freely behaving tadpoles."

> Results

> Page 4, line 6 "the stimuli we used were weaker" What is meant, the stimuli in the previous studies were weaker, or the stimuli in the present study? Maybe say the previous stimuli were very high contrast and did not reflect contrasts used in behavioral studies. Here, we presented a checkerboard"...

Changed to "the stimuli we used in the present study were weaker".

> Page 4 line 19 "spiky" again, imprecise

We like the word "spiky" (see above), as it captures the phenomelogical effect without making assuptions about underlying mechanisms. It is also intuitive and short, which aids understanding. If the reviewers insist, we can of course replace it with another term, but with this resubmission we hope to give it another try.

> Page 4, lines 21-24. I'm really confused about the degrees of freedom. Are you all loading in all observations from all cells? If so, the assumption of the statistical test that these are independent observations is violated. Many of these values were measured in the same cells with different current levels, so they are clearly not independent measurements. The analysis should essentially contain one observation per cell. (One "gain" value or number of spikes to some stimulus, etc..) Or could do a multi-variate ANOVA...

We did in fact run a multivariate linear model with repeated measures that may be described as multi-variate ANOVA; see responses above.

> Page 4: Lines 27, 28: This sentence is missing conjunctions "in order to map their amplitude transfer function" ?

Reworded: "We then mapped the amplitude tuning of neurons (also known as amplitude transfer function, or gain), by looking at how an increase in transmembrane conductance translated into increased spike output."

> Page 4: Lines 27, 28: What exactly is meant by the amplitude transfer function or amplitude tuning? Is there an equation?

At this point we describe tuning as the shape of an empirically observed curve (something that can be plotted, and compared from cell to cell), and do not yet use  equations to quantify it. Later in the text we parameterize these curves, to capture key differences between them, and provide all related equations.

> Page 4, lines 27 - 33: what figure panels are being described here?

We added a reference to Fig 1D,E.

> Page 6, lines 10-20. The authors point out that the amount of response adaptation in the conductance vs. number of spike curves varies as a function of the stimulus that the animal was exposed to. While it is true that this information can be gleaned by examining Figure 1E, the authors would be well served to make a second index, such as an adaptation index, in a panel F. This would really drive home the points the authors would like the reader to take away from the Figure. (I missed it the first time through, I just saw the reduced gain in Fig. 1E, not the adaptation.)

We introduce adaptation index (parametrized quantification of temporal tuning) later in the text.

> Page 6, lines 10-20. Need to cite Figure 1E

Done.

> Page 6, line 22: "homologous to superior colliculus"

Fixed.

> Page 7, line 3: after multi sensory stimulation, "tectal neurons were more excitable than after visual stimulation alone). Many issues. The average number of spikes is reported.. Is this across all conductance curve conditions or just one of them?

The averages were calculated across all testing conditions; we added this info to the sentence: "after four hours of multisensory stimulation, tectal neurons were more excitable than after visual stimulation alone (0.6$\pm$0.4 spikes, F(1,689)=11.2, p=8e-4; and 0.7$\pm$0.6 spikes, F(1,665)=41.7, p=2e-10, for sync and async respectively, across all testing conditions)."

> Next, in Fig 1E, "sync" and "flash" and "looming" all look pretty similar to each other.

We think that it rather reasonable that effects of 4 hours of stimulation, compared to control, are stonger than differences between different types of stimulation. There are differences however, and we explore this question in detail further in the text.

> What are the degrees of freedom? Again, they seem very high like each condition might be being treated as independent.

See other comments above. Generally, unless a mixed model is used, the second value for degrees of freedom, defining F-distributions, is close to the total number of measurements recorded in a study, even if a multivariate analysis of variance is used, and even for nested and non-orthogonal designs, such as blocking or repeated measures analysis. For our study in particular, the only way to bring the second value for degrees of freedom down, making it closer to the number of cells, rather than to the number of measurements, would be to use a mixed model. We used a mixed model (lmer) for a senstivity test, and observed that it provided qualitatively similar results, which is described in the method. Everywhere else we report the results from a fixed model, as it allows more straightforward interpretation, and, in our opinion, is more consistent with the experimental design.

> More stats: Are these multiple comparisons across conditions done in an unbiased manner, like with a Bonferroni correction or an ANOVA? It seems not.

1) As described above, we used a family of similar generalized linear models, followed by F-test for the analysis of variance, to compare across conditions. This type of analysis may be loosely referred to as ANOVA, although it is our understanding that most textbooks would not recommend referring our model as an "ANOVA", as it included two ordinal variables, and a factor to account for repeated measures.

2) It is true that in our analysis we compared several pairs of conditions: Flash to Control, Looming to Control, Flash to Looming, Sound to Control, Flash to Sync, and Flash to Async. However, all of these comparisons were easily interpretable, as they form a logical progression of stimuli. These planned comparisons consistuted a small share of all possible post-hoc comparisons between groups that one could have run on this data (6 out of 15 possible pairs). Typically, for planned comparisons, researchers do not use an adjustment for multiple comparisons.

3) Finally, in this paper we report exact p-values, instead of indicating whether they were above or below any arbitrary threshold. This gives the reader the freedom to interpret our data against any significance threshold of their choice. As expected, some p-values were larger, and some were smaller, but we are confident that our overall conclusions are reasonable, expecially if combining different pieces of evidence across the entire study.

Together, this makes us confident that a Bonferroni correction would be unnecessarily punitive for this type of analysis.

> Page 8, lines 2-12. "Average firing rate" is a more common term than "spikiness".

Please see our comments above. The term "average firing rate" is typically used to refer to actual average firing rates, either long-term or instaneous, in situations when relatively realistic neural activity is possible, such as in-vivo recordings, or during spontaneous activity. We however looked at a rather artificial excitability index, and we would prefer to stress that by naming this index, and defining it within the study.

> How about "gain" for amplitude tuning? I see that the authors are using the parameter a to indicate the non-linearity in the tuning curve. For "temporal tuning", why not use "adaptation index" or something like that, that could be plotted in Figure 1F? The "temporal tuning" is a measure of non-linearity, but it is not very intuitive. (It is not "incorrect" as the authors have it now, but I think it would be hard for the interdisciplinary readership of J Neuroscience to get on board easily.)

We would really prefer to retain terms "amplitude tuning" and "temporal tuning", as they better represent the main message of our paper, as we are trying to think of these values not in terms of abstract properties each neuron has, but as parameters that can tune a neuron to a certain type of synaptic inputs, in terms of their temporal properties, and their amplitude. We however now included the words "gain" and "adaptation index" in paragraphs that introduce values of amplitude and temporal tuning, respectively.

> Figure 2: Units needed for Y axis on A, B (units of a are spikes per conductance squared and units of amplitude tuning are spikes/conductance)

Both tuning parameters are dimensionless, as both conductance and amplitude were treated as rank-transformed, ordinal values (see methods).

> Page 8, about line 9: It is false that the numerical values are not interpretable, they are in an equation and have units.

We now replaced "not interpretable" with "not easily interpretable". We wanted our readers to know that a comparison of these values can be interpreted very easily, while exact numbers (say, 0.82) do not carry that much narrative weight. As mentioned above, they however do not have units (are dimensionless).

> Page 8, line 16-19: Cohen's D tells the discriminability rather than whether there is a difference between two quantities. A test like the Hoteling-2 test can provide significance measures for whether the means of the data in 2C are significant.

We initially reported Cohen's d values as a measure of effect sizes, not as a quantification of significance. Now we also ran Hotelling t-squared tests on 2D sets of points, representing amplitude and temporal tuning values for different cells, across different groups.

> Page 8, lines 26-35: Rather than only saying that groups that were significantly different in average values were also significantly different in variability, name the groups again. (Otherwise the reader has to jump all around.)

Fixed.

> Figure 2B and Figure 3A, 3D don't show any post-hoc comparisons among the groups. Which groups differ? I see it in the text but it is hard to take it home.

We added asterisks to mark significance compared to control. As not all pairwise comparisons make sense, we did not include brackets in the figure.

> Figure 3E and G: how can we reconcile these accounts? In each of the groups (3G), there is at best a very weak relationship between temporal tuning and synaptic duration. Somehow, if you take the means of all of these groups, you can fit a line through them (Figure 3E). Why is that the right thing to do? Aren't the field of individual points (partially-transparent background) and the data plotted in 3G the relative quantities?

We now added a paragraph describing this situation in detail, and reference a textbook name for this type of situation, when between-group and within-group analyses yield seemily opposite results (Simpson's paradox).

> Figure 3C shows that monosynaptic amplitude and synaptic duration are correlated, as one would expect from the definition of synaptic duration (center of mass). Basically, as the monosynaptic amplitude gets bigger, the center of mass will shift earlier and earlier. So what independent information is really being conveyed in the synaptic duration? The correlation in 3C is so strong, I don't see that it is truly an independent quantity that teaches us something beyond the monosynaptic input strength. The monosynaptic input strength is much more defined, why not use that for examining correlation with temporal tuning (or adaptation index), average firing rate, and gain. The "synaptic duration" isn't really a measure of synaptic duration, which makes Page 10, lines 3-23 really hard to interpret.

When comparing temporal properties of synaptic inputs to that of intrinsic tuning, it is important to make the two measures as similar (matching) as possible, as methodologically it is supposed to increase the power of our comparisons. In practice, if for intrinsic properties we quantify non-linearity of responses to currents of different duration, then we need to estimate current duration for our synaptic inputs. In other words, while synaptic current duration is largerly just a different way to quantify the relative impact of polysynaptic activity, it is a necessary step for a good analysis. We now reworded two sentences at the beginning of paragraph in question, to hopefully make this point more clear:

"To better match and compare temporal properties of synaptic inputs to those of intrinsic tuning, we calculated average "synaptic current duration" for each cell as a temporal "center of mass" of currents within the first 700 ms after optic chiasm stimulation (see "Methods"). Neurons with different contribution of early and late synaptic responses naturally had different synaptic current duration: cells with strong monosynaptic inputs had shorter currents, while polysynaptic activity made synaptic currents longer (Figure 3C; p=2e-16, r=$-$0.78, n=168)."

> Figure 3G, page 10 lines 24-40. Creation of "super groups" is totally bogus. This part should be dropped. I see no strong significant relationship between temporal tuning and synaptic duration, and am not convinced synaptic duration is a meaningful quantity.

We now removed this paragraph. In response to this comment, we initially attempted a different analysis of the interaction between synaptic and intrinsic properties across groups, but while some reasonable types of comparisons were significant, this analysis lacks power with our sample sizes, and also relies too much on post-hoc reasoning, so we decided to follow the advice and just remove this paragraph entirely. While we still suspect that stronger stimulation (in terms of its effect on average intrinsic properties) generally caused stronger disruption of the synaptic-intrinsic correlation, this statement is hard to fully justify with our current data.

> I am very confused by Figure 4. Is it supposed to tell us that average firing rate is correlated with sodium channel conductances? Why is this experiment being done? What do we learn about the brain? I mean this genuinely, I feel I do not understand what the authors are trying to say that is novel. Of course we would expect such a relationship. Why not show the 8 parameter model fit of the current tuning curves?

In the originally submitted version of this paper, we tried to show correlations betwee two estimations of spikiness (average spiking recorded in current and dynamic clamp modes) with Na conductances, to illustrate that one of the correlations is much stronger than the other. We agree however that showing the 8 parameter model fit is a better way of illustrating this point. We now show two scatterplots, for both dynamic and current clamp data, in which values predicted by the 8-parameter model are compared to actual measurements. It is clear to see that one of the models (that for current clamp data) works much better than the other one (for the dynamic clamp data).

> Page 11, "the mechanisms behind temporal intrinsic plasticity". I'm confused. This section and Figure 4 seems to focus on average firing rate / spikiness rather than the temporal tuning parameter. So why the heading?

Different fields have different criteria for what level of explanation is low enough to count as "mechanistic". We tried to predict cell spiking phenotypes, and responses to different simulated conductances, from their low-level electrophysiological properties. Had we succeeded in that, especially across treatment groups, we would have comfortably called it a mechanistic explanation. We show that for one type of measurements (current clamp), we have a tolerable level of mechanistic understanding of differences in observed cell excitability, while for more nuanced dynamic clamp experiments, we cound not predict neither average spiking, nor tuning parameters. In the figure 4, we mostly concentrate on average spiking, as it allows a visual comparison between the results obtained via two recording techniques.

> Discussion:

> Page 13, lines 18 - 24: the authors refer to the questions they asked in the introduction, without re-describing them. The reader cannot possibly remember and is forced to page back and forth. The writing must be so much better.

We reworded the beginning of the discussion, to make sure that we re-state all questions before providing the answers.

> "intrinsic plasticity does not simply adjust neuronal spikiness, but can regulate selectivity for inputs of different dynamics". How are the input dynamics different? I only see that monosynaptic magnitude is different across the conditions. The duration parameter does not reflect duration, just a weighted average of where the most input arrives, and is not related to the whole duration of synaptic input in a manner independent of the initial volley of input.)

The "synaptic duration" value is higher for those synaptic inputs that last longer (have a long polysyaptic tail after the original bout of monosynaptic currents), and is smaller for those inputs that are short (mostly limited to monosynaptic currents). Therefore, we believe that duration value does capture and reflect the duration. We chose the "weighted average" approach (the center of mass), as strictly speaking, in the presence of spontaneous activity there is no point when the synaptic resonse is truly over, and one has to go for a different way to quantify it.

While preparing the original version of this paper, we also tried a different way to estimate synaptic durations: we attempted to fit all synaptic currents with the same curve that we used for simulated conductances in our dynamic clamp experiments `I = g*(t/tau)*exp(1-t/tau)`. While this idea seemed good in theory, as it would have offered an "ideal" comparison with the dynamic clamp data, in practice synaptic curves were too variable in terms of their shape, decay, latency, and the amount of noise, making most curve fits really bad. This made us go for a simpler, robust to noise, and parsimonuous measure.

> "we show that synaptic and intrinsic temporal properties of neurons are homeostatically co-tuned" what does this mean?

We now made the Discussion section a bit longer, and so used longer, simpler statements, when describing the co-ordination of synaptic and intrinsic properties. For example, the sentece in question was replaced with this sentence: "In answer to our second question, about whether intrinsic and synaptic properties of tectal cells are in any way coordinated, here we show that intrinsic and synaptic temporal properties are co-tuned in the tectum, and moreover, that this co-tuning can be modified by sensory experience."

--------------------------

# Reviews from eLife, 2018 Aug 17 (rejection letter)

> ...While both BRE members found much to like about the manuscript, they found it dense going. But they both felt this paper belongs in a more specialized journal. A comment from one of them:

> The authors pose two questions in the introduction: 1. Is intrinsic plasticity limited to changes in excitability? 2. Whether intrinsic plasticity is a response to synaptic activation? There are data in other systems that make these questions "straw men". We know from the STG that intrinsic plasticity is not limited to change in spikiness alone and that there are several other parameters such as bursts, rebounds, spike widths etc which can and do get altered in the short and long term. There are also plenty of examples where one ion channel expression is co-regulated with another type of ion channel, or by slow action of growth factors, hormones and other neuromodulators. These are examples of intrinsic plasticity not governed by fast synaptic transmission alone.

This is of course true, but while many types of intricate effects of intrinsic plasticity are described in small networks, and especially in oscillatory networks, such as the STG, it is less common to consider these non-trivial effects in large feed-forward excitatory networks, such as networks in the cortex, or superior colliculus. It is natural for researchers to limit the complexity of their conceptual and computational models, and it is probably fair to say that as the number of neurons involved in a model increases, researchers often shift their attention from intrinsic mechanisms to connectomic approaches, approximating neural networks as a collection of nodes, linked with synapses of different strengths (e.g: Cline 2008; Kirby 2013). Even realistic biophysical models that represent neurons as multi-compartment systems with active dendritic currents often don't assume that these active currents may be modulated in response to activation (Moldwin 2018). When machine learning approaches are used as a model for brain development, intrinsic excitability is reflected in these models as a change in activation function of each artificial neuron (Marblestone 2016). This type of modeling implicitly assumes, however, that outside of obviously oscillatory networks (spinal cord, thalamus, etc.), intrisic excitability in biological networks is largely confined to overall changes in neuronal "responsiveness".

Here we show that even for large-scale feed-forward networks that do not seem to rely on oscillations, changes in intrinsic excitability go well beyond variations in "spikness". We think that it is an interesting observation, that may have far-reaching effects on network computations, and network development. We also suggest that similar effects of temporal tuning may be observed in other systems, such as mammalian cortex.

In the new version of this manuscript, we expanded the introduction, trying to make our message more clear, and also more nuanced.

&nbsp;

References:

Cline, H., & Haas, K. (2008). The regulation of dendritic arbor development and plasticity by glutamatergic synaptic input: a review of the synaptotrophic hypothesis. The Journal of physiology, 586(6), 1509-1517.

Kirkby, L. A., Sack, G. S., Firl, A., & Feller, M. B. (2013). A role for correlated spontaneous activity in the assembly of neural circuits. Neuron, 80(5), 1129-1144.

Marblestone, A. H., Wayne, G., & Kording, K. P. (2016). Toward an integration of deep learning and neuroscience. Frontiers in computational neuroscience, 10, 94.

Moldwin, T., & Segev, I. (2018). Perceptron learning and classification in a modeled cortical pyramidal cell. bioRxiv, 464826

> That aside, the authors make a big deal out of 'temporal tuning' but to me from fig 1 this looks like non-preference towards syn duration more than preference for any temporal window, as argued by the authors (compare fig 1E panels).

We think that a preference or non-preference for a certain synaptic duration can be described as a change in temporal selectivity. In the optic tectum, this temporal tuning is not as obvious as it would have been for an oscillatory network, as tectal neurons don't seem to express typical ionic currents associated with oscillatory networks (such as the h-current). Still, one can argue that it only makes our discovery more interesting: we show that neurons can change their temporal properties, even though they don't seem to be equipped with currents most researchers associate with temporal selectivity.

> There are other methodological issues as well: Did they randomize the order of injections?

No, the order of injections was always the same; it is now decribed in the "Methods": "Conductance curves were always presented in the order from the shortest to the longest, and this sequence was not randomized."

> Is a syn current that is more than 250 ms long 'monosynaptic' ?

We did not completely understand the question, but generally, for this model, monosynaptic component peaks at about 5-20 ms, and is over by 50-100 ms, as described in the "Methods" section.